Conversational expectations account for apparent limits on theory of mind use

### 2016-01-29 writing/cogsci-submission/cogsci_keysar.tex
Minor revisions to abstract

\begin{abstract}
People are capable of accurately reasoning about the mental states of other people using theory of mind. An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's perspective into account. This has been widely interpreted as a limitation on theory of mind use. In this paper we argue that these apparent failures are in fact successes. Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation. A listener who correctly reasons about the pragmatics of the task, thereby \emph{using} theory of mind, will rationally ignore whether an object is hidden to the speaker or not and will instead rely on the precision of the utterance to disambiguate. When an uncooperative confederate flouts these expectations, this strategy produces errors; when we allow participants to naturally interact, this strategy is justified and errors are reduced.
\ndg{make a bit less technical}

>>>>

\begin{abstract}
Theory of mind is a powerful cognitive ability: by the age of six, people are capable of accurately reasoning about the beliefs and desires underlying other people's behavior. An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's beliefs into account. This has been widely interpreted as evidence for limitations on theory of mind use. In this paper we argue that these apparent failures are in fact successes. Pragmatic language understanding requires sophisticated social reasoning about the speaker's goals and context. Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation. Listeners who are justified in expecting more informative utterances are thus tricked by confederates into making systematic patterns of errors. When we allow participants to naturally interact, their expectations are justified and errors are reduced.
\ndg{make a bit less technical}
